Conflict-Free Vertex Coloring Of Planar Graphs

### 2017-03-20 paper/paper.tex
Adding algorithm and chugging along

\documentclass{sig-alternate}
\usepackage{color}
\usepackage[colorinlistoftodos]{todonotes}

@@

\maketitle
\begin{abstract}
The conflict-free coloring problem is a variation of the vertex coloring problem, a classical NP-hard optimization problem. The conflict-free coloring problem aims to color the vertices of a graph in a way where every vertex is connected to at least one uniquely colored vertex.
\end{abstract}

@@

\subsection{Graph Coloring}
\label{sec:coloring}
A vertex coloring is an assignment of colors to each vertex of a graph $G$ such that no adjacent vertices share the same color. Mathematically, it can be described as a function $f : V \rightarrow S = \{1, 2, \dots, k\}$ such that $\forall u,w \in V$, if $(u,w) \in E$, then $f(u) \neq f(w)$. The \emph{chromatic number} of $G$, denoted $\chi(G)$, is the minimum number of colors, i.e. $|S|$, needed to properly color $G$. The \emph{vertex coloring} problem, when given a simple graph $G$, is to find $\chi(G)$. \cite{bondy1976graph}

@@

An example of a conflict-free coloring is shown in figure \ref{fig:cfcp-example}. It is worth noting that all proper vertex colorings are also conflict-free colorings. This is because in a proper vertex coloring, every vertex is its own conflict-free neighbor \cite{abel2017three}.

@@

\section{CF Coloring of General Graphs}

>>>>

\documentclass{sig-alternate}
\usepackage{color}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

@@

\maketitle
\begin{abstract}
The conflict-free coloring problem is a variation of the vertex coloring problem, a classical NP-hard optimization problem. The conflict-free coloring problem aims to color the vertices of a graph in a way where every vertex is connected to at least one uniquely colored vertex. This paper presents heuristics to solve the conflict-free coloring problem on both general graphs and planar graphs. This paper also looks into bounds on the number of colors needed.
\end{abstract}

@@

\subsection{Graph Coloring}
\label{sec:coloring}
A vertex coloring is an assignment of colors to each vertex of a graph $G$ such that no adjacent vertices share the same color. Mathematically, it can be described as a function $f : V \rightarrow S = \{1, 2, \dots, k\}$ such that $\forall u,w \in V$, if $(u,w) \in E$, then $f(u) \neq f(w)$. A vertex coloring that satisfies this constraint is called a \emph{proper coloring}. The \emph{chromatic number} of $G$, denoted $\chi(G)$, is the minimum number of colors, i.e. $|S|$, needed to properly color $G$. The \emph{vertex coloring} problem, when given a simple graph $G$, is to find $\chi(G)$. \cite{bondy1976graph}

@@

An example of a conflict-free coloring is shown in figure \ref{fig:cfcp-example}. It is worth noting that all proper vertex colorings are also conflict-free colorings. This is because in a proper vertex coloring, every vertex is its own conflict-free neighbor \cite{abel2017three}. We observe from figures \ref{fig:vcp-example} and \ref{fig:cfcp-example}, $\chi(H) = 3$ and $\chi_{CF}(H) = 2$.

The domination number of $G$, denoted $\gamma(G)$, is the size of a minimum dominating set of G. A dominating set is a subset $D$ of $V$ such that every vertex not in $D$ is adjacent to at least one vertex in $D$. The \emph{conflict-free domination number} for some $k$, denoted $\gamma_{CF}^k(G)$, is the minimum number of vertices that have to be colored in a conflict-free k-coloring of $G$. An example of a dominating set is $\{3, 5\}$ of graph $H$ where $\gamma(H) = 2$.

@@

\section{CF Coloring of General Graphs}
For the rest of this paper, we will consider the NP-complete problem of conflict-free k-colorability. Given a graph $G$ and an integer $k$, determine if graph $G$ can be colored using $k$ or fewer colors. Although all graphs can be conflict-free colored, a given graph $G$ may not be able to be colored with $k$ colors.

As shown with graph $H$ in section \ref{sec:background}, even though all proper vertex colorings are also conflict-free colorings, fewer colors can usually be used. This leads us to use heuristics that differ from the vertex coloring problem. Abel et al. \cite{abel2017three} present an efficient heuristic to color certain general graphs with $k$ colors in a conflict-free manner. They call this heuristic \emph{iterated elimination of distance-3-sets}.

\subsection{Guaranteeing CF k-Colorability}
It is wasteful to spend time coloring a graph that cannot be conflict-free colored with $k$ colors. This leads Abel et al. \cite{abel2017three} to provide sufficient criterion to guarantee the conflict-free k-colorability of a graph. To introduce this criterion, we need to define a few more terms.

\subsection{Iterated Elimination of Distance-3-Sets}

\begin{algorithm}
\caption{Iterated elimination of distance-3-sets} \label{alg:elimination}
\begin{algorithmic}[1]
\State $i \gets 1$, $\chi \gets \emptyset$
\State Remove all isolated paths from $G$
\While{$G$ is not empty}
	\State $D \gets \emptyset$
	\ForAll{components of $G$}
		\State Pick a vertex $v$ and add it to $D$
		\While{$\exists u$ at distance $\geq 3$ $\forall v \in D$}
			\State Pick $u$ at distance 3 from some vertex in $D$
			\State $D \gets D \cup \{ w \}$
		\EndWhile
		\ForAll{$u \in D$}
			\State Color $u$ with color $i$
		\EndFor
		\State $i \gets i + 1$
		\State Remove $N[D]$ from $G$
		\State Remove all isolated paths from G
	\EndFor
\EndWhile
\State Color all removed isolated paths using color $i$
\end{algorithmic}
\end{algorithm}
